A More Complete Algebra for File Synchronization

### 2015-11-10 2015/paper.tex
Introduced existential extensions to the algebra; finished proof

\usepackage{alltt}
\usepackage{comment}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{setspace}
\usepackage{fullpage}
\usepackage{graphicx}
\usepackage{afterpage}
\usepackage{centernot}

@@

\newcommand{\cbrk}{\mathtt{break}}
\newcommand{\fscommand}[2]{{#1#2}}
\newcommand{\fsregcommandchar}[1]{\mathtt{#1}}
\newcommand{\fsregcommand}[2]{\fscommand{\fsregcommandchar{#1}}{\fsregcommandchar{#2}}}
\newcommand{\cbb}{\fsregcommand{\empt}{\empt}}
\newcommand{\cbf}{\fsregcommand{\empt}{F}}
\newcommand{\cbd}{\fsregcommand{\empt}{D}}
\newcommand{\cfb}{\fsregcommand{F}{\empt}}
\newcommand{\cff}{\fsregcommand{F}{F}}
\newcommand{\cfd}{\fsregcommand{F}{D}}
\newcommand{\cdb}{\fsregcommand{D}{\empt}}
\newcommand{\cdf}{\fsregcommand{D}{F}}
\newcommand{\cdd}{\fsregcommand{D}{D}}
\newcommand{\cxy}{\fscommand{X}{Y}}
\newcommand{\cyz}{\fscommand{Y}{Z}}
\newcommand{\cxz}{\fscommand{X}{Z}}
\newcommand{\czw}{\fscommand{Z}{W}}
\newcommand{\cqr}{\fscommand{Q}{R}}

\newcommand{\eqext}{\sqsubseteq}
\newcommand{\eqnrw}{\sqsupseteq}
\newcommand{\nequiv}{\not\equiv}

@@

@@

Similarly to \cite{NREC:alg}, we have two kinds of judgements
for sequences \(S_1\) and \(S_2\):
\begin{itemize}
\item \(S_1\equiv S_2\), or \emph{\(S_1\) is algebraically equivalent to
\(S_2\)}. Its intended interpretation is that they act in the same way on all
filesystems.
\item \(S_1\eqext S_2\), or \emph{\(S_2\) extends \(S_1\)}; its intended
interpretation is that they act in the same way on all
filesystems that \(S_1\) does not break.
\end{itemize}

@@

By a {\bf minimal sequence or set} of commands we mean a sequence / set in which
there is at most one command on each path.

@@

In each component that is larger than one, the paths are directly related
and the commands form construction or destruction chains.
To order the commands,
select the command with the topmost path, so that there would be no command
on its parent path. If the command removes content at this path, the command
needs to follow all other commands in the component, and
the component is a destruction component.
Otherwise the command with the topmost component needs to precede other commands
in the component, and it is a construction component.
Repeat the above process of splitting commands into subsets on the remaining commands
in the component. Recursively, this way we can determine all allowed orders of the commands.

We use $\orderset{U}$ to denote the set of sequences that contain all commands in $U$
and represent all allowed orders. As a specific case, we can regard a sequence of commands $S$
as a set and determine its $\orderset{S}$.
At times by $\orderset{U}$ we mean an arbitrary sequence in the set.

% (O-D)
{\bf Lemma.} If $U$ contains commands on both $p$ and $\pp$, then
there is a $U^*\in\orderset{U}$ where they are next to each other.

{\bf Lemma.} For any minimal sequence $S$ not equivalent to $\cbrk$,
$\orderset{S} = \extset{S}$.

@@

{\bf Lemma.} \((\forall\FS: S\FS = \empt) \Rightarrow S \equiv \cbrk\)

@@

@@

If $A$ and $B$ are minimal sequences and there is at least one filesystem neither of them breaks,
we write $A\circ B$. In this case based on the soundness of the algebra
we also know that $A\nequiv\cbrk$ and $B\nequiv\cbrk$.
Also, if there are commands on path $p$ in both $A$ ($\cxy(p, V_Y)$) and $B$ ($\czw(p, V_W)$)
then their input types must match ($X=Z$).

{\bf Lemma.} Let $A$ and $B$ be two minimal sequences so that $A\circ B$.
Then their intersection can also be applied to any filesystem
either can be applied to without breaking it.

Proof.
First we prove the following.
Mark all commands in $A$ that also appear in $B$.
Then $\exists A^* \in \orderset{A}$ in which the marked commands are at the beginning.

@@

If $B$ has a command on $q$, let it be $\cqr(q, V_R)$. As $A\circ B$, we know
that $Q=Z$. From 
O-D (TODO ref)
we also know that there is a $B^*\equiv B$ where $\cxy$ and $\cqr$ are next to each other,
and from the construction and destruction laws we know that $\cxy$ determines that like
$\czw(q, V_W)$ in $A$, any command on $q$ must precede it in $B$. As $B$ is minimal, $B^*$
contains the same commands, and so it is the $\cqr$ command that precedes $\cxy$ in $B^*$.
Going back to the construction and destruction laws we see that the output type of the first command
is always the same, and it is either $D$ or $\empt$. Therefore $R=W$ and $\cqr(q, V_R)=\czw(q, V_W)$
as $|\setd|=|\setb|=1$. This is a contradiction as $\czw$ was not marked, but we see it must also be in $B$.

If $B$ has no command on $q$,

>>>>

\usepackage{alltt}
\usepackage{comment}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{setspace}
\usepackage{fullpage}
\usepackage{graphicx}
\usepackage{afterpage}
\usepackage{centernot}

@@

\newcommand{\cbrk}{\mathtt{break}}
\newcommand{\fscommand}[2]{{#1#2}}
\newcommand{\fsregcommandchar}[1]{\mathtt{#1}}
\newcommand{\fsregcommand}[2]{\fscommand{\fsregcommandchar{#1}}{\fsregcommandchar{#2}}}
\newcommand{\cbb}{\fsregcommand{\empt}{\empt}}
\newcommand{\cbf}{\fsregcommand{\empt}{F}}
\newcommand{\cbd}{\fsregcommand{\empt}{D}}
\newcommand{\cfb}{\fsregcommand{F}{\empt}}
\newcommand{\cff}{\fsregcommand{F}{F}}
\newcommand{\cfd}{\fsregcommand{F}{D}}
\newcommand{\cdb}{\fsregcommand{D}{\empt}}
\newcommand{\cdf}{\fsregcommand{D}{F}}
\newcommand{\cdd}{\fsregcommand{D}{D}}
\newcommand{\cxy}{\fscommand{X}{Y}}
\newcommand{\cyz}{\fscommand{Y}{Z}}
\newcommand{\cxz}{\fscommand{X}{Z}}
\newcommand{\czw}{\fscommand{Z}{W}}
\newcommand{\cqr}{\fscommand{Q}{R}}
\newcommand{\cqq}{\fscommand{Q}{Q}}

\newcommand{\eqext}{\sqsubseteq}
\newcommand{\eqnrw}{\sqsupseteq}
\newcommand{\nequiv}{\not\equiv}
\newcommand{\wrext}{\stackrel{\mathtt{br}}{\sqsubseteq}}
\newcommand{\wrnrw}{\stackrel{\mathtt{br}}{\sqsupseteq}}
\newcommand{\coworks}{\circ}
\newcommand{\ncoworks}{\centernot\circ}

@@

\theoremstyle{definition}
\newtheorem{mydef}{Definition}

\newtheorem{mylem}{Lemma}

@@

Similarly to \cite{NREC:alg}, we have two kinds of judgements
for sequences \(S_1\) and \(S_2\):
\begin{mydef}
\(S_1\equiv S_2\), or \emph{\(S_1\) is algebraically equivalent to
\(S_2\)}. Its intended interpretation is that they act in the same way on all
filesystems. It is a commutative and transitive relation.
\end{mydef}
\begin{mydef}
\(S_1\eqext S_2\), or \emph{\(S_2\) extends \(S_1\)}; its intended
interpretation is that they act in the same way on all
filesystems that \(S_1\) does not break. It is a transitive relation.
\end{mydef}

@@

\begin{mydef}{Minimal set / sequence}
By a minimal sequence or set of commands we mean a sequence / set in which
there is at most one command on each path.
\end{mydef}

@@

In each component that is larger than one, the paths are directly related
and the commands form construction or destruction chains.
To order the commands,
select the command with the topmost path, so that there would be no command
on its parent path. If the command removes content at this path, then it
needs to follow all other commands in the component, and
the component is a destruction component.
Otherwise the command with the topmost component needs to precede other commands
in the component, and it is a construction component.
Repeat the process of splitting commands into subsets on the remaining commands
in the component. Recursively, this way we can determine all allowed orders of the commands.

\begin{mydef}{$\orderset{U}$}
We use $\orderset{U}$ to denote the set of sequences that contain all commands in $U$
and represent all allowed orders.
\end{mydef}

As a specific case, we can regard a sequence of commands $S$
as a set and determine its $\orderset{S}$.
At times by $\orderset{U}$ we mean an arbitrary sequence in the set.

\begin{mylem}\label{lemma:neighbour}
If $U$ contains commands on both $p$ and $\pp$, then
there is a $U^*\in\orderset{U}$ where they are next to each other.
\end{mylem}

\begin{mydef}
$\extset{S} := \{S^*|S\eqext S^*\}$
\end{mydef}

\begin{mylem}\label{lemma:minextset}
For any minimal sequence $S$ not equivalent to $\cbrk$, $\extset{S} = \{S^*|S\equiv S^*\}$.
\end{mylem}
\begin{proof}
This is because as $S$ has at most one command on each path, only equivalence axioms
can be applied to it.
\end{proof}

\begin{mylem}\label{lemma:minextorder}
For any minimal sequence $S$ not equivalent to $\cbrk$,
$\orderset{S} = \extset{S}$.
\end{mylem}

@@

\begin{mylem}
\((\forall\FS: S\FS = \empt) \Rightarrow S \equiv \cbrk\)
\end{mylem}

@@

\section{Extending the algebra}

We have seen that commands can break a filesystem
either because their input type does not match the filesystem at the given path,
or because the filesystem loses its tree property.

\begin{mydef}{Fully typed sequences}
We call a sequence of commands fully typed if whenever it breaks a filesystem
it is explained by a mismatch on the input types of the commands it contains.
TODO: define fully typed sequences in terms of the algebra?
\end{mydef}

To aid the construction of fully typed sequences, we reintroduce
two assertion commands that either break the filesystem
or leave it unchanged: $\cbb(p)$ and $\cdd(p)$.
As $|\setb|=|\setd|=1$, we do not note the exact output value.
These can be used to test whether $\FS(p)=\empt$ or $\FS(p)=D$.

Furthermore, we note the following axioms:

\begin{itemize}
\item Laws for construction typing:
   \begin{gather*}
            [\cbf(p, F)] \equiv [\cdd(\pp); \cbf(p, F)]\\
            [\cbd(p, D)] \equiv [\cdd(\pp); \cbd(p, D)]
   \end{gather*}
\item Laws for destruction typing:
   \begin{gather*}
            [\cdb(\pp)] \equiv [\cbb(p); \cdb(\pp)]\\
            [\cdf(\pp, F)] \equiv [\cbb(p); \cdf(\pp, F)]
   \end{gather*}
\end{itemize}

Separately, we extend the algebra with the following judgements and axioms:

\begin{mydef}{$A\coworks B$}
or the domains of $A$ and $B$ overlap.
Its intended interpretation is that
there is at least one filesystem neither $A$ nor $B$ breaks.
It is a commutative (but not transitive) relation.
For $A\ncoworks B$ we say that $A$ and $B$ are disjoint.
\end{mydef}

\begin{mydef}{$A\wrext B$}
or the domain of $B$ extends that of $A$.
Its intended interpretation is that
$B$ does not break any filesystem $A$ does not break.
It is a transitive relation.
\end{mydef}

\begin{itemize}
\item $A\coworks B \Rightarrow A\nequiv\cbrk$
\item $A\eqext B \Rightarrow A\wrext B$
\item $A\coworks B\wrext C \Rightarrow A\coworks C$
\item $A\ncoworks B\wrnrw C \Rightarrow A\ncoworks C$
\item $A\ncoworks B \Rightarrow A;S\ncoworks B$
\end{itemize}

Axioms:

\begin{itemize}
\item $\cxy(p,V_Y)\ncoworks \czw(p,V_W)$ where $X\neq Z$
% \item Disjoint laws for construction:
%    \begin{gather*}
%    % bd>bf
%             [\cbd(\pp, D); \cbf(p, F)] \ncoworks [\cbf(p, F)]\\
%    % bd>bd
%             [\cbd(\pp, D_1); \cbd(p, D_2)] \ncoworks [\cbd(p, D_2)]\\
%    % fd>bf
%             [\cfd(\pp, D); \cbf(p, F)] \ncoworks [\cbf(p, F)]\\
%    % fd>bd
%             [\cfd(\pp, D_1); \cbd(p, D_2)] \ncoworks [\cbd(p, D_2)]
%    \end{gather*}
% \item Disjoint laws for destruction:
%    \begin{gather*}
%    % fb<db ??
%             [\cfb(p); \cdb(\pp)] \ncoworks [\cdb(\pp)]\\
%    % fb<df ??
%             [\cfb(p); \cdf(\pp, F)] \ncoworks [\cdf(\pp, F)]\\
%    % db<db ??
%             [\cdb(p); \cdb(\pp)] \ncoworks [\cdb(\pp)]\\
%    % db<df ??
%             [\cdb(p); \cdf(\pp, F)] \ncoworks [\cdf(\pp, F)]
%    \end{gather*}
\end{itemize}

\begin{mylem}
If $A$ and $B$ are minimal sequences, $A\coworks B$,
and there are commands on path $p$ in both $A$ ($\cxy(p, V_Y)$) and $B$ ($\czw(p, V_W)$)
then their input types must match ($X=Z$).
TODO: Can this be derived from the algebra?
\end{mylem}

\begin{mylem}
If $A$ is fully typed and $S;A$ is minimal, then $S;A\wrext A$
TODO: Is this an axiom?
\end{mylem}

\begin{mylem}\label{lemma:ncowprefix}
$A\ncoworks B \Rightarrow S;A\ncoworks B$ if $A$ is fully typed and $S;A$ is minimal.
\end{mylem}
\begin{proof}
This follows from above.
\end{proof}

@@

\begin{mylem}
Let $A$ and $B$ be two minimal sequences so that $A\coworks B$.
Then their intersection can also be applied to any filesystem
either can be applied to without breaking it.
\end{mylem}

\begin{proof}
First we prove the following.
Mark all commands in $A$ that also appear in $B$.
Then $\exists A^* \in \orderset{A}$ in which the marked commands are at the beginning.

\medskip

@@

If $B$ has a command on $q$, let it be $\cqr(q, V_R)$. As $A\coworks B$, we know
that $Q=Z$. From lemmas \ref{lemma:neighbour}, \ref{lemma:minextset} and \ref{lemma:minextorder}
we also know that there is a $B^*\equiv B$ where $\cxy$ and $\cqr$ are next to each other,
and from the construction and destruction laws we know that $\cxy$ determines that like
$\czw(q, V_W)$ in $A$, any command on $q$ must precede it in $B$. As $B$ is minimal, $B^*$
contains the same commands, and so it is the $\cqr$ command that precedes $\cxy$ in $B^*$.
Going back to the construction and destruction laws we see that the output type of the first command
is always the same, and it is either $D$ or $\empt$. Therefore $R=W$ and $\cqr(q, V_R)=\czw(q, V_W)$
as $|\setd|=|\setb|=1$. This is a contradiction as $\czw$ was not marked, but we see it must also be in $B$.

If $B$ has no command on $q$, $A$ and $B$ cannot work on the same filesystem,
as the second commands in all construction and destruction laws require a different type of input
as the first commands. This is a contradiction as we supposed $A\coworks B$.

In terms of the algebra, this latter argument could be rephrased as follows.
Split $A$ into $[A_0;\czw(q, V_W);\cxy(p, V_Y);A_1]$.
Add command $\cqq(q)$ before $\cxy(p, V_Y)$ to $B$ according to the construction or destruction typing laws.
This yields $B^*\equiv B$, which we split into $[B_0;\cqq(q);\cxy(p, V_Y);B_1]$.
We know $[\czw(q, V_W);\cxy(p, V_Y)]$ is fully typed (TODO: why?)
and $A$ is minimal, so $[A_0;\czw(q, V_W);\cxy(p, V_Y)]$ is minimal.
We also know $[\cqq(q);\cxy(p, V_Y)]$ is fully typed and $B^*$ is minimal, so $[B_0;\cqq(q);\cxy(p, V_Y)]$ is minimal.
But we know $[\czw(q, V_W);\cxy(p, V_Y)]\ncoworks [\cqq(q);\cxy(p, V_Y)]$
as comparing the construction and destruction laws to their typing counterparts we know that $Z\neq Q$,
and so from lemma \ref{lemma:ncowprefix} and $A\ncoworks B \Rightarrow A;S\ncoworks B$
we know that $A\ncoworks B^*\equiv B$ and so $A\ncoworks B$ which is a contradiction.
(TODO: Is there a point in proving this this way?)

\medskip

\end{proof}
